\systemnamenospace: Head Orientation-Based Selection\\in Physical Spaces

### 2014-08-14 doc/sui14/applications.tex
rewrote sentence about sample tasks
Grammar changes and comments
bjoern's editing pass

Our prototype, \changes{(based on {\em Naive IR}, \footnote{For this particular configuration, the environment is not dense enough, and {\em Naive IR} performs well for head orientation based targeting proof-of-concept.})} includes three smart devices: a lamp and a fan that could be switched on and off; a smart TV with playback, volume and navigation controls (see Figure~\ref{fig:smart-home}). The on/off appliances are using relays, and the smart TV is a 30\inch{} display connected to a laptop. User interfaces for each were pre-defined in our application.

%
We set the devices up in a simulated living room environment and invited 14 users to step through a predefined set of tasks to control the appliances at a distance. \changes{These tasks include turning off the light, playing movie, turning up the volume, etc., which are some sample tasks that people would normally do at home.}

>>>>

Our prototype, includes three smart devices: a lamp and a fan that could be switched on and off; and a smart TV with playback, volume and navigation controls (see Figure~\ref{fig:smart-home}). \changes{The prototype used \em{Naive IR}, as the small number of target devices made selection without disambiguation possible.} The appliances are switched with 120V AC relays. The smart TV is a 30\inch{} display connected to a laptop. User interfaces for each were pre-defined in our application.

%
We set the devices up in a simulated living room environment and invited 14 users to step through a predefined set of tasks to control the appliances at a distance. \changes{The tasks included turning off a lamp, playing a movie and turning up the volume.}

### 2014-08-14 doc/sui14/iteration1.tex
bjoern's editing pass

%!TEX root = uist14.tex

@@

\subsection{Technique}
In our first implementation, we use IR for scanning as described in the previous section. For the refinement stage, we simply show a list of the subset of targets that have received IR signals on the Glass display. Users swipe to select from that list and tap to confirm the intended target.

@@

\subsection{Method}
We deployed 10 wireless nodes in an indoor environment (Figure~\ref{fig:targeting-study-layout}), each with a number ID and a letter representing the name. For the study, we recruited \changes{16} participants from our institution \changes{(8 males and 6 females)} \changes{using email invitations}. \changes{They are composed of undergrad students, graduate students, and university staff. Their background ranges from Information(6), Engineering(4), Math and Science(4), Design(1), Others(1)}  \changes{14} had never used Google Glass before, so we offered a tutorial before the experiment in order to introduce the device. Four participants wore prescription glasses, which makes Glass more cumbersome to use and adjust and may have affected their task performance. In this study, the IR LED was fixed and not repositionable as shown in Figure~\ref{fig:glass}.

\begin{figure}[t]
\centering
\includegraphics[width=0.9\columnwidth]{figures/study-layout1.png}
\caption{In the targeting study, participants were asked to find and select one of 10 targets in the lab environment.}
\label{fig:targeting-study-layout}
\end{figure}

In the within-subject study, half of the participants performed {\em IR selection} first and the other half used {\em list selection} first. For each selection condition, we conducted 15 target selections by randomly choosing from all the targets. During the study, we measured the {\bf target acquisition time} for each target selection. \changes{At the beginning of each acquisition, participants were asked to stand at a fixed position approximately 10-12 feet away from the nodes, looking down until the target is provided. As they attempt to connect, they may move around if they decided it would help them select. After each task, they were asked to move back to the starting position.} Afterward, participants were asked to complete a survey of primarily open-ended \changes{questions} about their experience.

%% remove this for now
%% Our measurements suggest that IR communication can be targeted to an area about 60-120cm in diameter, up to 5m in front of the user. These values are a reasonable match for selecting targets in a room-size environment. A wider beam would lead to an increased chance of multiple appliances receiving IR signals simultaneously. A narrower beam will make targeting more challenging, given the precision constraints of human head movement.

\subsection{Results}
Our results indicate that {\em IR selection} outperforms {\em list selection}.  The average target acquisition time for {\em IR selection} is 6.67 seconds 
while {\em list selection} took 8.86 seconds (see Figure~\ref{fig:ir_vs_list}). A t-test shows a significant difference ($t(279)=-3.81, \changes{p < 0.001}$).

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.95\columnwidth]{figures/result_study1b.pdf}
\caption{Boxplot and CDF of target acquisition time when refinement is needed or not in {\em IR selection}}.
\label{fig:with_without_refinement}
\end{figure}

@@

To understand how scanning and refinement contribute to total selection time, we split the data from {\em IR selection} into two parts -- trials that required refinement and ones that did not. It takes 6.40 seconds (on average) to complete  a selection without any refinement, but 9.16 seconds with refinement, indicating that an additional 2.76 seconds are needed for disambiguation (see Figure~\ref{fig:with_without_refinement}). This difference is significant $(t(19)=-2.7827, p=0.012)$ using t-test.
%  \bjoern{I don't understand why we have 279 DF in the previous test but only 19 here??}). 
%% \ben{Because the times when refinement is needed is only 18 times, so the DF is bounded by the samples.}
Because many targets were spaced far apart, refinements were only necessary in 10\% of total {\em IR selection} trials in this study.

To further generalize the results, in Figure~\ref{fig:time-vs-list-order}, we show the acquisition time for each individual target in {\em list selection} condition, ordered by their relative position in the list. Since with IR technique, the acquisition time is invariant from each target's order, we use solid lines to represent the average performance of {\em IR selection}. From this figure, we can see that once there are more than 6 targets, the average acquisition time will be larger than IR selection, even if disambiguation is needed. %Therefore, we argue that once there are more targets in the physical space, the {\em list selection} method is not scalable. Further more, in real-life scenarios, targets would have longer names than an alphabet letter. This would increase the searching and text comparison time and also reduce the number of targets being shown in the UI screen.

As for the qualitative feedback, 11 out of \changes{16} participants preferred IR selection over list mode (\changes{3} preferred list, \changes{2} was undecided). \changes{When asked: \studyquote{When you have multiple candidate targets within sight, which method for selecting between them did you use?} (multiple choices), among 16 participates, 81\% chose {\em adjust head so that only one candidate is signaled}, 56\% chose {\em tap (to enter refinement stage) and then choose between list}, and 25\% chose {\em walk closer to the candidate for a better result}. Although through our observation, many participants walked one or two steps during some tasks, this suggests that participants viewed their moving as a followed-through action of adjusting their heads instead of an intention to walk closer to the targets.

>>>>

%!TEX root = sui14.tex

@@

\newpage
\subsection{Technique}
In our first implementation, we use IR for scanning as described in the previous section. For the refinement stage, we simply show a list of the subset of targets that have received IR signals on the Glass display. Users swipe to select from that list and tap to confirm the intended target.

@@

\subsection{Method}
We deployed 10 wireless nodes in an indoor environment (Figure~\ref{fig:targeting-study-layout}), each with a number ID and a letter representing the name. For the study, \changes{we recruited 16 participants from our institution (8 males and 6 females) by email.} \changes{Participants included undergraduate and graduate students, as well as university staff. Their educational backgrounds included Information Science (6), Engineering (4), Math and Science (4), Design (1), Others(1).}  \changes{14} had never used Google Glass before, so we offered a tutorial before the experiment in order to introduce the device. Four participants wore prescription glasses, which makes Glass more cumbersome to use and adjust and may have affected their task performance. In this study, the IR LED was fixed and not repositionable as shown in Figure~\ref{fig:glass}.

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/study-layout1.png}
\caption{In the targeting study, participants were asked to find and select one of 10 targets in the lab environment.}
\label{fig:targeting-study-layout}
\end{figure}

In the within-subject study, half of the participants performed {\em IR selection} first and the other half used {\em list selection} first. For each selection condition, we conducted 15 target selections by randomly choosing from all the targets. During the study, we measured the {\bf target acquisition time} for each target selection. \changes{At the beginning of each acquisition, participants were asked to stand at a fixed position approximately 10-12 feet away from the nodes, looking down until a target is announced. Participants were allowed to physically move towards targets if they decided that this would help them with the task. After each task, they were asked to move back to the starting position.} Afterwards, participants were asked to complete a survey of primarily open-ended \changes{questions} about their experience.

%% remove this for now
%% Our measurements suggest that IR communication can be targeted to an area about 60-120cm in diameter, up to 5m in front of the user. These values are a reasonable match for selecting targets in a room-size environment. A wider beam would lead to an increased chance of multiple appliances receiving IR signals simultaneously. A narrower beam will make targeting more challenging, given the precision constraints of human head movement.
\newpage %bjoern - hack fix to avoid orphaned heading
\subsection{Results}
Our results indicate that {\em IR selection} outperforms {\em list selection}.  The average target acquisition time for {\em IR selection} is 6.67 seconds 
while {\em list selection} took 8.86 seconds (see Figure~\ref{fig:ir_vs_list}). A t-test shows a significant difference ($t(279)=-3.81, \changes{p < 0.001}$).

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.95\columnwidth]{figures/result_study1b.pdf}
\caption{Boxplot and CDF of target acquisition time when refinement is needed or not in {\em IR selection}.}
\label{fig:with_without_refinement}
\end{figure}

@@

To understand how scanning and refinement contribute to total selection time, we split the data from {\em IR selection} into two parts -- trials that required refinement and ones that did not. It takes 6.40 seconds (on average) to complete  a selection without any refinement, but 9.16 seconds with refinement, indicating that an additional 2.76 seconds are needed for disambiguation (see Figure~\ref{fig:with_without_refinement}). This difference is significant $(t(19)=-2.7827, \changes{p<0.05})$.% using t-test. %p=0.012
%  \bjoern{I don't understand why we have 279 DF in the previous test but only 19 here??}). 
%% \ben{Because the times when refinement is needed is only 18 times, so the DF is bounded by the samples.}
Because many targets were spaced far apart, refinements were only necessary in 10\% of total {\em IR selection} trials in this study.

To further generalize the results, in Figure~\ref{fig:time-vs-list-order}, we show the acquisition time for each individual target in the {\em list selection} condition, ordered by their relative position in the list. Since with IR technique, the acquisition time is invariant from each target's order, we use solid lines to represent the average performance of {\em IR selection}. From this figure, we can see that once there are more than 6 targets, the average acquisition time will be larger than IR selection, even if disambiguation is needed. %Therefore, we argue that once there are more targets in the physical space, the {\em list selection} method is not scalable. Further more, in real-life scenarios, targets would have longer names than an alphabet letter. This would increase the searching and text comparison time and also reduce the number of targets being shown in the UI screen.

\changes{In the elicited qualitative feedback, 11 out of 16 participants preferred IR selection over list mode (3 preferred list, 2 were undecided)}. \changes{When asked: \studyquote{When you have multiple candidate targets within sight, which method for selecting between them did you use?}, 81\% chose {\em adjust head so that only one candidate is signaled}, 56\% chose {\em tap (to enter refinement stage) and then choose between list}, and 25\% chose {\em walk closer to the candidate for a better result} (multiple choices were allowed)}. \changes{While we observed more participants walking one or two steps during some tasks, this suggests that participants viewed their moving as a followed-through action of adjusting their heads instead of an intentional to walk closer to the targets.}

### 2014-08-14 doc/sui14/iteration2.tex
bjoern's editing pass

We have empirically measured the intensity distribution at the receiver for this configuration in Figure~\ref{fig:measurement}. Our measurements confirm that angular difference has a large effect on the intensity readings (see how the IR intensity distribution changes when the \changes{angle has increased}) \achal{this sounds awkward to me - maybe "Note the changes in IR intensity distribution when the angle increases}.

\begin{figure}[t]
\centering
\includegraphics[width=0.95\columnwidth]{figures/IRIntensityDistribution.pdf}
\caption{This figure shows an empirical measurement of IR intensity at different positions. We measure the intensity at a plane every meter away from the IR emitter. On that plane, a sample is taken every 5 centimeters, \changes{and we stop when the reading is smaller than ambient noise.} \changes{To match our assumption about the relationship between angles and intensities, the vertical axis is changed to the degree of the angle.} The size and color brightness represent the intensity of the readings for visualization.}
\label{fig:measurement}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.95\columnwidth]{figures/study-layout2.pdf}
\caption{The environment setup for our second study. In comparison to our first study, we have deliberately increased the target density.}
\label{fig:study-layout2}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.95\columnwidth]{figures/result_study2.pdf}
\caption{Boxplot and CDF of the target acquisition time in {\em Naive IR} and {\em Intensity IR} conditions.}
\label{fig:study2}
\end{figure}

\sean{If possible, add how we observed participants physical movements in this iteration. Better if we can compare to i1.} One side effect that we have observed in this approach is that the {\em Intensity IR} sometimes eliminates the desired target during the {\em scanning} stage. Out of 300 trials, \changes{the target was accidentally eliminated in 13 (4.3\%).} This is higher than the error rate for {\em Naive IR} (5 out of 300, ~1.6\%). \changes{Even though the higher error rate increases the chance of multiple attempts in target selection, our analysis above illustrates that overall performance of {\em Intensity IR} is still better than {\em Naive IR}. }  % but this is the trade-off for the overall faster acquisition time.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End:

>>>>

We have empirically measured the intensity distribution at the receiver for this configuration in Figure~\ref{fig:measurement}. \changes{Our measurements confirm that angular difference has a large effect on the intensity readings, with rapid fall-off at increasing angles.}
%(see how the IR intensity distribution changes when the \changes{angle has increased}) \achal{this sounds awkward to me - maybe "Note the changes in IR intensity distribution when the angle increases}.

\begin{figure}[t]
\centering
\includegraphics[width=0.95\columnwidth]{figures/IRIntensityDistribution.pdf}
\caption{\changes{Empirical measurement of IR intensity at different positions. We measure the intensity on a horizontal plane at one-meter intervals from the IR emitter.} On that plane, a sample is taken every 5 centimeters, \changes{and we stop when the reading is smaller than ambient noise.} \changes{To match our assumption about the relationship between angles and intensities, the vertical axis is changed to the degree of the angle.} The size and color brightness represent the intensity of the readings for visualization.}
\label{fig:measurement}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/study-layout2.pdf}
\caption{The environment setup for our second study. In comparison to our first study, we have deliberately increased the target density.}
\label{fig:study-layout2}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/result_study2.pdf}
\caption{Boxplot and CDF of the target acquisition time in {\em Naive IR} and {\em Intensity IR} conditions.}
\label{fig:study2}
\end{figure}

\sean{If possible, add how we observed participants physical movements in this iteration. Better if we can compare to i1.} One side effect that we have observed in this approach is that the {\em Intensity IR} sometimes eliminates the desired target during the {\em scanning} stage. Out of 300 trials, \changes{the target was accidentally eliminated in 13 (4.3\%).} This is higher than the error rate for {\em Naive IR} (5 out of 300, ~1.6\%). \changes{Even though the higher error rate increases the chance of multiple attempts in target selection, our analysis above illustrates that overall performance of {\em Intensity IR} is still better than {\em Naive IR}. }  % but this is the trade-off for the overall faster acquisition time.
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End:

### 2014-08-14 doc/sui14/iteration3.tex
bjoern's editing pass

%!TEX root = uist14.tex
\section{Iteration 3: Head motion refinement}
\label{sec:iteration-3:-head}

@@

In the learning stage, as the user scans over targets in their normal
use, the system attains the absolute orientation of each device from
IR and motion sensors. From this information it can abstract out
\changes{the} relative positions \changes{of targets} and construct an
{\em adjacency map}. The absolute orientations in the map cannot be
applied to all indoor environments, since the user's movements through
the space could change the relationships between targets. However,
with the constraint that the targets are spread around the periphery
\changes{at similar distances}, their relative orientations are stable
(see Figure~\ref{fig:third_principle}). And since we are using IR at
the same time, the learning process is transparent to the user and the
map can be built without an explicit calibration phase.

@@

\changes{After the map is created, the user can enter a quasi-mode for
refinement by holding down on the touchpad. In this quasi-mode, one device
lights up at a time. When the user turns his head in the direction of another
device, the light switches to that device. Therefore, the user can move between
devices one at a time with slight head movements. This prediction is
implemented by first calculating the user's direction of motion, and searching
through the adjacency map for the nearest device in that direction.  To
calculate the direction, we maintain a circular buffer of the last 10 sensor
measurements, passed through a low pass filter. The direction is then
calculated as the difference between the last sample and the first sample.
Note that we use a hysteresis method to temporarily lock to the current device
- if the variance of the buffer is below a threshold, we assume no movement has
occurred and do not select another device.}

\subsection{Evaluation}
We evaluate the head motion refinement method by holding an informal study and
collecting qualitative feedback from a subset of 4 users from the Iteration 2
study. In the study, we asked users to try cycling through the targets using
the new quasi-mode.

The users had strong preferences for the new method of refinement.
\changes{Though we lack formal data to evaluate the performance of
this method, from our observation during the experiments, we found
that each trial became much easier than previous iterations. We
conducted a survey to collect qualitative feedback after the
experiment.} On a scale from 1-7, 1 being the least mental effort and
7 being the most mental effort users rated the old technique 4.25 and
the new technique 2 on average. 100\% indicated a preference for head
movement to list navigation. One user referenced the issue of naming
targets that the list necessitates, preferring the experience of
\studyquote{matching visual cues rather than numbers}.  Another
participant remarked that it \studyquote{just made more sense} and was
a \studyquote{more natural way for demonstrating intentionality}. The
users preferred the new mapping in relation to the whole environment:
\studyquote{it leveraged the spatial sense that I already had just by
using the system}. They were also delighted to avoid list navigation,
which they now called \studyquote{difficult} and \studyquote{painful}.

>>>>

%!TEX root = sui14.tex
\newpage
\section{Iteration 3: Head motion refinement}
\label{sec:iteration-3:-head}

@@

In the learning stage, as the user scans over targets in their normal
use, the system attains the absolute orientation of each device from
IR and motion sensors. From this information it can abstract out
\changes{the} relative positions \changes{of targets} and construct an
{\em adjacency map}. The absolute orientations in the map cannot be
applied to all indoor environments, since the user's movements through
the space could change the relationships between targets. However,
with the constraint that the targets are spread around the periphery
\changes{at similar distances}, their relative orientations are stable
(see Figure~\ref{fig:third_principle}). \changes{This learning process can be transparent to the user and the
map can be built without an explicit calibration phase.}

@@

\changes{After the map is created, the user can enter a quasi-mode for
refinement by holding down on the touchpad. In this quasi-mode, a single selected device
lights up at a time. When the user turns his head in the direction of another
device, the selection (and indicator light) switches to that device. Therefore, the user can move between
devices one at a time with slight head movements. This prediction is
implemented by first calculating the user's direction of head motion, and searching
through the adjacency map for the nearest device in that direction.  To
calculate the direction, we maintain a circular buffer of the last 10 sensor
measurements, passed through a low pass filter. The direction is then
calculated as the difference between the last sample and the first sample.
We use a hysteresis method to avoid spurrious selection changes
--- if the variance of the buffer is below a threshold, we assume no movement has
occurred and do not select another device.}

\subsection{Evaluation}
We evaluated the head motion refinement method through an informal study and
collected qualitative feedback from a subset of 4 users from the Iteration 2
study. In this evaluation, we asked users to cycle through multiple targets using
the new quasi-mode.

The users had strong preferences for the new method of refinement.
\changes{Our observations suggest that each trial became much easier than previous iterations. We
conducted a survey to collect qualitative feedback after the
experiment.} 
On a scale from 1-7, 1 being the least mental effort and
7 being the most mental effort users rated the old technique 4.25 and
the new technique 2 on average. All users indicated a preference for head
movement to list navigation. One user referenced the issue of naming
targets that the list necessitates, preferring the experience of
\studyquote{matching visual cues rather than numbers}.  Another
participant remarked that it \studyquote{just made more sense} and was
a \studyquote{more natural way for demonstrating intentionality}. The
users preferred the new mapping in relation to the whole environment:
\studyquote{it leveraged the spatial sense that I already had just by
using the system}. They were also delighted to avoid list navigation,
which they now called \studyquote{difficult} and \studyquote{painful}.
